---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
  instalar la última versión desde github de spark
---


```{r}
install.packages("RSQLite")
install.packages("ggthemes")
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

//esto es hive

src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10")
class(dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10"))
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, "SELECT * FROM orders_tbl LIMIT 100")
orders_prev
```
 
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}

products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently?
* The most popular hour of doing order? 
* The most reordered products?
* The less reorderd product?
* Plot histogram of number of products in an order.
* What is the most popular department? 

## Whisker, magrittr

```{r}
dbGetQuery(sc, "select * from orders_tbl limit 5")
```

```{r}

# * Which day of week people order most fequently?
dbGetQuery(sc, "select a.order_dow, a.freq from  (SELECT order_dow, count(order_dow) as freq FROM orders_tbl group by order_dow ) a ORDER BY a.freq DESC LIMIT 1")

```
```{r}
# * The most popular hour of doing order? 

dbGetQuery(sc, "DESC orders_tbl")

dbGetQuery(sc, "select a.order_hour_of_day, a.freq from  
           (SELECT order_hour_of_day, count(order_hour_of_day) as freq FROM orders_tbl group by order_hour_of_day ) a 
           ORDER BY a.freq DESC LIMIT 1")

```

```{r}
#* The most reordered products?
# [1] "order_products__prior_tbl" "order_products__train_tbl" "orders_tbl"   "products_tbl"     
dbGetQuery(sc, "DESC order_products__prior_tbl")
dbGetQuery(sc,"select * from order_products__prior_tbl limit 5")

dbGetQuery(sc,"
          SELECT x.product_id from
           (select a.product_id as product_id, a.freq from 
                (select product_id, sum(reordered) as freq from 
                      order_products__prior_tbl group by product_id) a ORDER BY FREQ DESC LIMIT 10 ) x")
dbGetQuery(sc,"select * from products_tbl where product_id=24852")


dbGetQuery(sc,"
          SELECT x.*, y.product_name from
           (select a.product_id as product_id, a.freq from 
                (select product_id, sum(reordered) as freq from 
                      order_products__prior_tbl where reordered>0  group by product_id) a ORDER BY FREQ desc LIMIT 1 ) x
          inner join (select * from products_tbl) as y on x.product_id=y.product_id")


```

```{r}
# * Plot histogram of number of products in an order.

dbGetQuery(sc, "DESC order_products__prior_tbl")
dbGetQuery(sc, "select * from order_products__prior_tbl limit 10")


dbGetQuery(sc, "  ( select order_id, product_id, count(1) as freq from order_products__prior_tbl group by order_id,product_id limit 10) ")


# de las dos formas peta por memoria

dbGetQuery(sc, "select a.order_id, sum(a.freq) from 
                  ( select order_id, product_id, count(1) as freq from order_products__prior_tbl group by order_id,product_id limit 10) a
                  group by a.order_id")

dbGetQuery(sc, "select order_id, count(1) as product_id from order_products__prior_tbl group by order_id limit 10")

 
 
```

```{r]
* What is the most popular department? 
```

```{r}
# * The less reorderd product?

dbGetQuery(sc,"
          SELECT x.product_id from
           (select a.product_id as product_id, a.freq from 
                (select product_id, sum(reordered) as freq from 
                      order_products__prior_tbl group by product_id) a ORDER BY FREQ ASC LIMIT 10 ) x")
dbGetQuery(sc,"select * from products_tbl where product_id=36380")

dbGetQuery(sc,"
          SELECT x.*, y.product_name from
           (select a.product_id as product_id, a.freq from 
                (select product_id, sum(reordered) as freq from 
                      order_products__prior_tbl where reordered>0 group by product_id ) a ORDER BY FREQ ASC LIMIT 1 ) x
          inner join (select * from products_tbl) as y on x.product_id=y.product_id")


```

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```
```{r}
product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, plot the most popular product (with templating department).
* For given day, rank department popularity.

## More on orders (Window Functions)



```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```
```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```

