---
title: "Instacart Exploratory Analysis"
output:
  html_document: default
  html_notebook: default
  instalar la última versión desde github de spark
---


```{r}
# install.packages("RSQLite")
# install.packages("ggthemes")
source("readDataToMemory.R")
readInstacart()

library(DBI)
library(ggplot2)
library(ggthemes)

# //esto es hive

src_tbls(sc)
```

### Order products

`order_products__*.csv`

These files specify which products were purchased in each order.

```{r}
dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10")
class(dbGetQuery(sc, "SELECT * FROM order_products__prior_tbl LIMIT 10"))
```
 
## orders

`orders.csv`

This file tells to which set (prior, train, test) an order belongs. 


```{r}
orders_prev <- dbGetQuery(sc, "SELECT * FROM orders_tbl LIMIT 100")
orders_prev
```
 
```{r}
dbGetQuery(
  sc,
  "
SELECT COUNT(DISTINCT product_id) AS products
FROM order_products__prior_tbl
  "
  )
```

```{r}
products_popularity_txt <- "
SELECT product_id
,   COUNT(1) AS n_orders
FROM order_products__prior_tbl
GROUP BY product_id
ORDER BY n_orders DESC
LIMIT 50
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Products

`products.csv`

```{r}
dbGetQuery(sc, "SELECT * FROM products_tbl LIMIT 20")
```

## Joins

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

```{r}

products_popularity %>%
  ggplot(
    aes(reorder(product_name, n_orders, function(x) x), 
        n_orders)) +
  geom_bar(stat="identity", fill='darkorange') +
  coord_flip() + 
  scale_y_continuous(label=scales::comma) +
  xlab("product") +
  ylab("Number of orders") +
  theme_minimal()
```

```{r}
products_popularity_txt <- "
SELECT op.product_id
,   n_orders
,   product_name
FROM (
    SELECT product_id
    ,   COUNT(1) AS n_orders
    FROM order_products__prior_tbl
    GROUP BY product_id
    ORDER BY n_orders DESC
    LIMIT 30) op
LEFT JOIN (
    SELECT product_id
    ,   product_name
    FROM products_tbl) p
ON op.product_id = p.product_id
"

products_popularity <-
  dbGetQuery(sc, products_popularity_txt)

products_popularity
```

## Questions

* Which day of week people order most fequently?
* The most popular hour of doing order? 
* The most reordered products?
* The less reorderd product?
* Plot histogram of number of products in an order.
* What is the most popular department? 

## Whisker, magrittr

```{r}
dbGetQuery(sc, "select * from orders_tbl limit 5")
```

```{r}

# * Which day of week people order most fequently?
dbGetQuery(sc, "select a.order_dow, a.freq from  (SELECT order_dow, count(order_dow) as freq FROM orders_tbl group by order_dow ) a ORDER BY a.freq DESC LIMIT 1")

```
```{r}
# * The most popular hour of doing order? 

dbGetQuery(sc, "DESC orders_tbl")

dbGetQuery(sc, "select a.order_hour_of_day, a.freq from  
           (SELECT order_hour_of_day, count(order_hour_of_day) as freq FROM orders_tbl group by order_hour_of_day ) a 
           ORDER BY a.freq DESC LIMIT 1")

```

```{r}
#* The most reordered products?
# [1] "order_products__prior_tbl" "order_products__train_tbl" "orders_tbl"   "products_tbl"     
dbGetQuery(sc, "DESC order_products__prior_tbl")
dbGetQuery(sc,"select * from order_products__prior_tbl limit 5")

dbGetQuery(sc,"
          SELECT x.product_id from
           (select a.product_id as product_id, a.freq from 
                (select product_id, sum(reordered) as freq from 
                      order_products__prior_tbl group by product_id) a ORDER BY FREQ DESC LIMIT 10 ) x")
dbGetQuery(sc,"select * from products_tbl where product_id=24852")


dbGetQuery(sc,"
          SELECT x.*, y.product_name from
           (select a.product_id as product_id, a.freq from 
                (select product_id, sum(reordered) as freq from 
                      order_products__prior_tbl where reordered>0  group by product_id) a ORDER BY FREQ desc LIMIT 1 ) x
          inner join (select * from products_tbl) as y on x.product_id=y.product_id")


```

```{r}
# * Plot histogram of number of products in an order.

dbGetQuery(sc, "DESC order_products__prior_tbl")
dbGetQuery(sc, "select * from order_products__prior_tbl limit 10")


dbGetQuery(sc, "  ( select order_id, product_id, count(1) as freq from order_products__prior_tbl group by order_id,product_id limit 10) ")


# de las dos formas peta por memoria

dbGetQuery(sc, "select a.order_id, sum(a.freq) from 
                  ( select order_id, product_id, count(1) as freq from order_products__prior_tbl group by order_id,product_id limit 10) a
                  group by a.order_id")

dbGetQuery(sc, "select order_id, count(1) as product_id from order_products__prior_tbl group by order_id limit 10")

 
 
```


```{r}
dbGetQuery(sc,
           "select  n_products, count(distinct order_id) as n_orders  from ( select order_id, count(1) as n_products from order_products__prior_tbl group by order_id) op
           group by n_products") %>% ggplot(aes(n_products, n_orders)) + geom_bar(stat='identity')+ theme_minimal() 

dbGetQuery(sc,
           "select  n_products, count(distinct order_id) as n_orders  from ( select order_id, count(1) as n_products from order_products__prior_tbl group by order_id) op
           group by n_products having n_orders>1000") %>% ggplot(aes(n_products, n_orders)) + geom_col(fill="darkorange")+ theme_minimal() 

order_products__prior %>%
  group_by(order_id) %>% summarise(n_products=n()) %>% group_by(n_products) %>% filter(n_products<=50) %>%
  summarise(n_orders=n()) %>% collect()%>% ggplot(aes(n_products, n_orders)) + geom_col(fill="darkorange")+ theme_minimal() 

```

```{r]
* What is the most popular department? 
```

```{r}
# * The less reorderd product?

dbGetQuery(sc,"
          SELECT x.product_id from
           (select a.product_id as product_id, a.freq from 
                (select product_id, sum(reordered) as freq from 
                      order_products__prior_tbl group by product_id) a ORDER BY FREQ ASC LIMIT 10 ) x")
dbGetQuery(sc,"select * from products_tbl where product_id=36380")

dbGetQuery(sc,"
          SELECT x.*, y.product_name from
           (select a.product_id as product_id, a.freq from 
                (select product_id, sum(reordered) as freq from 
                      order_products__prior_tbl where reordered>0 group by product_id ) a ORDER BY FREQ ASC LIMIT 1 ) x
          inner join (select * from products_tbl) as y on x.product_id=y.product_id")


```

```{r}
library(whisker)
library(magrittr)

product_times_ordered_txt <- "
SELECT n_orders
,   COUNT(DISTINCT user_id) AS n_users
FROM (
    SELECT o.user_id
    ,   COUNT(DISTINCT op.order_id) AS n_orders
    FROM (
        SELECT order_id
        FROM order_products__prior_tbl
        WHERE product_id = {{productId}}
    ) op
    LEFT JOIN (
        SELECT  user_id
        ,   order_id
        FROM orders_tbl) o
    ON op.order_id = o.order_id
    GROUP BY o.user_id) opo
GROUP BY n_orders
ORDER BY n_orders
"

data <- list(productId = 24852)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  cat
```
```{r}

data <- list(productId = 1)

product_times_ordered_txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) %>%
  ggplot(aes(n_orders, n_users)) + 
  geom_line() +
  theme_minimal()
```

## Questions

* For given department, give a list of th 10  the most popular product (with templating department).
* For given day of the week, rank department popularity.

## More on orders (Window Functions)


```{r}
dbGetQuery(sc,"SELECT * FROM ORDER_PRODUCTS__PRIOR_TBL LIMIT 10")
dbGetQuery(sc,"SELECT * FROM ORDERS_TBL LIMIT 10")
dbGetQuery(sc,"SELECT * FROM PRODUCTS_TBL LIMIT 10")

```


```{r}
# * For given department, give a list of th 10  the most popular product (with templating department).

# dbGetQuery(sc,"
txt<-          "SELECT x.*, y.product_name,y.department_id from
           (select a.product_id as product_id, a.freq from 
                (select product_id, sum(reordered) as freq from 
                      order_products__prior_tbl where reordered>0 group by product_id ) a ORDER BY FREQ DESC ) x
          inner join (select * from products_tbl) as y on x.product_id=y.product_id WHERE y.department_id= {{depid}} limit 10"

data <- list(depid = 19)

txt %>%
  whisker.render(data) %>%
  cat

txt %>%
  whisker.render(data) %>%
  dbGetQuery(sc, .) 

```

```{r}
dbGetQuery(sc,"SELECT * FROM ORDER_PRODUCTS__PRIOR_TBL LIMIT 10")
dbGetQuery(sc,"SELECT * FROM ORDERS_TBL LIMIT 10")
dbGetQuery(sc,"SELECT * FROM PRODUCTS_TBL LIMIT 10")

```

```{r}
# * For given day of the week, rank department popularity.

# dbGetQuery(sc,"
txt<-          "select a.* from order_products__prior_tbl as a inner join orders_tbl as b inner join products_tbl as c 
                on a.order_id=b.order_id and a.product_id=c.product_id limit 10"

txt%>%dbGetQuery(sc, .) 
# data <- list(depid = 19)
# 
# txt %>%
#   whisker.render(data) %>%
#   cat
# 
# txt %>%
#   whisker.render(data) %>%
#   dbGetQuery(sc, .) 

```

```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```

```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```








## More on orders (Window Functions)



```{r}
dbGetQuery(sc, 
"
SELECT days_since_prior_order
,   COUNT(1) AS N
FROM orders_tbl
GROUP BY days_since_prior_order
") 
```


```{r}
dbGetQuery(sc, 
"SELECT * 
FROM orders_tbl
WHERE user_id <= 10
")
```
```{r}
dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```
* Product that is usually orders in numerous baskets

```{r}
"
SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC
LIMIT 100
" %>%
dbGetQuery(sc, .)
```

```{r}

# could modify the query and add product name?
"
select a.*, b.product_name  from (SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number DESC) as a 
inner join 
(select * from products_tbl) as b on a.product_id=b.product_id
LIMIT 100
" %>%
dbGetQuery(sc, .) 
```


```{r}

# products that is usually orders in small baskets
"
select a.*, b.product_name  from (SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number ASC) as a 
LEFT join 
(select * from products_tbl) as b on a.product_id=b.product_id
LIMIT 100
" %>%
dbGetQuery(sc, .) 
```
```{r}

dbGetQuery(sc, 
"SELECT * FROM ORDERS_TBL LIMIT 10")


dbGetQuery(sc, 
"SELECT * FROM order_products__train_tbl limit 100")


dbGetQuery(sc, 
"SELECT user_id
,   order_id
,   FIRST_VALUE(order_id)  OVER(partition by user_id ORDER BY order_number) AS first_order
FROM orders_tbl
WHERE user_id <= 10 
")
```


* Could modify the query and add prduct name?

```{r]}
# For 10 first users (in all their orders) find days since first order

"
select a.*, b.product_name  from (SELECT product_id
,   AVG(products_number) AS avg_products_number
FROM (
    SELECT order_id
    ,   product_id
    ,   COUNT(product_id)  OVER(partition by order_id) AS products_number
    FROM order_products__train_tbl
) pn
GROUP BY product_id
ORDER BY avg_products_number ASC) as a 
LEFT join 
(select * from products_tbl) as b on a.product_id=b.product_id
LIMIT 100
" %>%
dbGetQuery(sc, .) 
```


```{r}

# * Could modify the query and add product name?

# * Product that is usually orders in small baskets.

# * For 10 first users and all their orders find days since first order.
```

```{r}
orders %>%
  filter(user_id <= 10) %>%
  select(user_id, order_number, days_since_prior_order) %>%
  arrange(user_id, order_number) %>%
  mutate(days_since_first_order = cumsum(days_since_prior_order)) %>%
  mutate(days_since_first_order=ifelse(is.null(days_since_prior_order), 0, days_since_first_order))
```


```{r}
# https://cwiki.apache.org/confluence/display/Hive/LanguageManual+WindowingAndAnalytics


"SELECT 
user_id
, days_since_prior_order
, sum(days_since_prior_order) 
OVER (PARTITION BY user_id ORDER BY order_number ROWS BETWEEN UNBOUNDED PRECEDING AND CURRENT ROW) as days_since_first_order, order_number
FROM orders_tbl where 
user_id<=10 
order by user_id, 
order_number" %>% dbGetQuery(sc,.)

"SELECT 
user_id
, days_since_prior_order
, COALESCE( sum(days_since_prior_order) 
OVER (PARTITION BY user_id ORDER BY order_number ROWS BETWEEN UNBOUNDED PRECEDING AND CURRENT ROW),0) as days_since_first_order, order_number
FROM orders_tbl where 
user_id<=10 
order by user_id, 
order_number" %>% dbGetQuery(sc,.)

"SELECT 
user_id
, days_since_prior_order
, COALESCE( sum(days_since_prior_order) 
OVER (PARTITION BY user_id ORDER BY order_number ),0) as days_since_first_order, order_number
FROM orders_tbl where 
user_id<=10 
order by user_id, 
order_number" %>% dbGetQuery(sc,.)

``` 